---
title: "A workflow for model-averaged renewal-process inference with marp"
author: "marp authors"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A workflow for model-averaged renewal-process inference with marp}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE
)
```

## Overview

The `marp` package fits six parametric renewal-process models to positive
inter-event times: Poisson (exponential waiting times), Gamma, log-logistic,
Weibull, log-normal, and Brownian passage time (BPT). It compares the candidate
models using AIC and calculates AIC-weighted model-averaged estimates.

The main fitted quantities are the mean inter-event time, the cumulative event
probability at a user-supplied time `y` stored on the **logit scale**, and the
hazard evaluated at user-supplied times `t` stored on the **log scale**.

This vignette demonstrates the standard fitted-object workflow. The final
confidence-interval step is shown but not evaluated because the existing
studentized procedure uses nested bootstrap sampling and is intentionally
computationally expensive.

## Prepare input data

`marp()` expects a numeric vector of positive inter-event times. The following
small reproducible example is simulated from a Gamma distribution.

```{r data}
library(marp)

set.seed(42)
dat <- rgamma(50, shape = 3, rate = 0.01)
summary(dat)
```

The remaining inputs control the displayed estimates. `t` gives the times for
log-hazard evaluation, `y` gives the time for the logit event probability, and
`m` controls repeated random-start optimizations in candidate models that use
`nlm()`.

The reference/generating-model codes are 1 Poisson, 2 Gamma, 3 log-logistic,
4 Weibull, 5 log-normal, and 6 BPT. A reference model is most naturally known
in a simulation; here code 2 matches the distribution used to generate `dat`.

```{r inputs}
t <- seq(100, 200, by = 20)
y <- 304
m <- 3
```

## Fit and inspect the candidate models

The primary interface fits all six candidates, selects the lowest-AIC model,
and calculates AIC-weighted estimates.

```{r fit}
set.seed(42)
fit <- marp(dat, t, m, y, which.model = 2)
fit
```

Printing the fitted object gives a concise model-comparison table, AIC weights,
the selected model, and the principal model-averaged estimates. A structured
summary provides the model-specific parameters and fit criteria.

```{r summarize}
fit_summary <- summary(fit)
fit_summary
```

The six rows are always in the documented candidate-model order. Parameters
are model-specific, so they should be interpreted using the documentation for
the corresponding model-specific fitter. For example, the Gamma parameters
are shape and rate, whereas the Weibull parameters are scale and shape.

## Inspect selected and model-averaged quantities

The S3 print and summary methods are the usual entry points. Existing named
list components remain available for programmatic use and backward
compatibility.

```{r extract}
fit$weights_AIC
fit$model_best
fit$mu_best
fit$mu_aic
fit$pr_aic
fit$haz_aic
```

`mu_best` is the mean estimate from the lowest-AIC model, while `mu_aic` is the
AIC-weighted mean estimate. `pr_aic` is an AIC-weighted average on the **logit
event-probability scale**. `haz_aic` contains AIC-weighted values on the
**log-hazard scale**, with one value for every element of `t`.

The supplied reference model is also retained. These quantities are useful for
simulation comparisons and do not imply that the true model is known in an
observational application.

```{r reference}
fit$mu_gen
fit$pr_gen
fit$haz_gen
```

## Confidence intervals

The standard `confint()` method delegates to the package's existing
`marp_confint()` implementation. Because the fitted object deliberately does
not store the original observations, the original data and bootstrap sizes
must be supplied explicitly.

```{r confidence-intervals, eval=FALSE}
set.seed(42)
ci <- confint(
  fit,
  data = dat,
  B = 99,
  BB = 99,
  level = 0.95
)
ci
```

The public function can equivalently be called directly with significance
level `alpha = 1 - level`:

```{r direct-confidence-intervals, eval=FALSE}
ci_direct <- marp_confint(
  data = dat,
  m = m,
  t = t,
  B = 99,
  BB = 99,
  alpha = 0.05,
  y = y,
  which.model = 2
)
```

Both forms use the same numerical engine. The studentized interval calculation
generates `BB` double-bootstrap samples for each of `B` bootstrap samples and
fits multiple renewal models repeatedly. Production runs can therefore take
substantially longer than the point-estimation workflow. Set a random seed for
reproducibility and choose `B` and `BB` appropriate to the analysis; the values
shown above illustrate the API rather than prescribing settings for every use.

## Summary

A typical analysis uses `marp()` followed by `print()` and `summary()` to
compare candidate models and inspect selected and model-averaged estimates.
Named components support programmatic extraction, while `confint()` provides a
conventional interface to the existing bootstrap confidence-interval engine.
